---
title: 'Create Blood Bowl roster plots'
author: "Gertjan Verhoeven"
date: '2023-01-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r warning=FALSE, message=FALSE}
# Load packages
library(tidyverse)
library(ggfortify)
library(ggrepel)
library(readODS)
```


# Read and prep the data

We start with reading in the scraped FUMBBL roster data, this is generated by the `fumbbl_rosters` Jupyter Notebook.

```{r}
df_rosters1 <- read.csv(file = "input/df_rosters_road_to_malta.csv")

df_rosters1$group_name <- "Road to Malta"

group_name <- "Road to Malta"

df_rosters <- df_rosters1 %>%
  filter(position != "") %>%
  mutate(cnt = 1) %>%
  rename(skill_name = name)
```

```{r}
# add xtra stuff
extra_stuff <- unique(df_rosters %>% 
                    mutate(apothecary = ifelse(apothecary == "Yes", 1, 0)) %>%
  select(group_name, team_id, roster.name, coach_name, rerolls, assistantCoaches, cheerleaders, apothecary) %>%
    pivot_longer(cols = -c(group_name, team_id, roster.name, coach_name), values_to = "cnt") %>%
    filter(cnt > 0) %>%
    mutate(position = name) %>%
    select(group_name, team_id, roster.name, coach_name, position, cnt) %>%
    mutate(player_id = 99999999) %>%
    mutate(number = 99) %>%
    mutate(skill_id = NA) %>%
    mutate(name = NA)) %>%
    rename(skill_name = name)

df_rosters <- df_rosters %>%
  select(-c(X, rerolls, assistantCoaches, cheerleaders, apothecary))

df_rosters <- rbind(df_rosters, extra_stuff)


```

# filter out BB2016 teams (Bretonnian)


```{r}
df_rosters <- df_rosters %>%
  filter(roster.name != "Bretonnian")
```

# Label old and new amazon, and old and new norse rosters

amazon team with old/new mix 5141 roster. roster contains positionals from old and new!!!

```{r}
# drop old amazon team in WC training
df_rosters <- df_rosters %>%
  filter(team_id != 836802)
```


```{r}
 old_amazon_team_ids <- df_rosters %>% filter(position == "Koka Kalim Blitzer") %>%
  filter(roster.name != "Amazon Eurobowl 2022") %>%
  group_by(team_id) %>%
  summarise(n()) %>%
  pull(team_id)
```   

```{r}
df_rosters <- df_rosters %>% 
  mutate(roster.name = ifelse(team_id %in% old_amazon_team_ids, "Amazon Team of Legend", roster.name))
```

There are also rosters with Norse Catcher AND Norse Raider. (old/new) march 2022, roster id 5144. This roster contains 

```{r}
 old_norse_team_ids <- df_rosters %>% filter(position %in% c("Norse Thrower", "Norse Catcher", "Norse Lineman")) %>%
  filter(roster.name != "Norse Eurobowl 2022") %>%
  group_by(team_id) %>%
  summarise(n()) %>%
  pull(team_id)
```

```{r}
df_rosters <- df_rosters %>% 
  mutate(roster.name = ifelse(team_id %in% old_norse_team_ids, "Norse Teams of Legend", roster.name))
```


```{r}
df_rosters %>% 
  filter(grepl("Amazon", roster.name)) %>%
  group_by(roster.name, position) %>%
  summarise(n = n())
```

# Check rosters without skills

```{r}
df_rosters %>%
  mutate(w_skill = ifelse(skill_name %in% c("", NA), 0, 1)) %>%
  group_by(group_name, team_id) %>%
  summarise(n_skills = sum(w_skill)) %>%
  mutate(has_skills = ifelse(n_skills == 0, 0, 1)) %>%
  group_by(group_name) %>%
  summarise(perc = mean(has_skills))
```
So the superleague has missing skills.
Hypothesis : teams that only played in round 0 Q2 2022 have missing skills?

# Drop teams without skills.

```{r}
teams_to_drop <- df_rosters %>%
  mutate(w_skill = ifelse(skill_name %in% c("", NA), 0, 1)) %>%
  group_by(group_name, team_id) %>%
  summarise(n_skills = sum(w_skill)) %>%
  filter(n_skills == 0) %>%
  pull(team_id)

```
```{r}
df_rosters <- df_rosters %>%
  filter(!(team_id %in% teams_to_drop))
```


# Cleanup the skill names

```{r}
df_rosters <- df_rosters %>%
  mutate(skill_name = str_replace_all(skill_name, 
            pattern = "^_", replacement = "")) %>%
  mutate(skill_name = replace_na(skill_name, ""))
```


# condense skill stacking rows

A player with two skills has two rows, collapse into one, concatenating the two skills.
We sort by skill_id to avoid mirror combinations.

```{r}
df_rosters <- df_rosters %>%
  arrange(player_id, -skill_id) %>%
  group_by(across(c(-skill_id, -skill_name))) %>%
  summarise(skill_name = paste(skill_name, collapse = ""))

```
# wrangle bribes into cnt 0-3 bribes

Cannot easily seperate out the GEt the ref bribes from the  roster bribes

```{r}
df_rosters %>% 
  ungroup() %>%
  filter(grepl("bribe", position)) %>%
  distinct(position) %>%
  pull()

```

```{r}
df_rosters <- df_rosters %>%
  mutate(cnt = ifelse(grepl("bribe", position), as.integer(substr(position, 1, 2)), cnt)) %>%
  mutate(position = ifelse(grepl("bribes", position), "1 bribe", position)) 
```


# Add skill colors and cost (metadata)

```{r}
library(bbrosterplots)
df_rosters <- df_rosters %>% select(-player_id) %>% ungroup() # PM BUGFIX

df_rosters <- bbrosterplots::add_roster_metadata(df_rosters)
```


<!--
# create color table

```{r eval = FALSE}
skill_colors <- df_rosters %>%
  group_by(skill_name) %>%
  summarise(n = n())

#write_ods(skill_colors, "skill_colors.ods")
```


# Add skill color table

```{r}
skill_colors <- read_ods("230218 bb_skill_colors.ods")

skill_colors <- skill_colors %>%
  mutate(skill_name = replace_na(skill_name, "")) %>%
  select(-n)
```


# create code list with roster cost

create empty list with all roster positions for each race
```{r}

out <- df_rosters %>%
  group_by(position, roster.name) %>%
  summarise(n = n())

```
manually fill in cost and sort order

# add codelist with roster cost

```{r}
cost <- read_ods("230217_bb_rosters_cost.ods")

cost <- cost %>%
  select(-n)
```

# check for missings

check if dataset has new positionals or star players etc.

```{r}
out <- out %>% 
  left_join(cost, by = c("position", "roster.name"))
```

```{r, eval = FALSE}
write_ods(out, "bb_rosters.ods")
```


```{r}
df_rosters <- df_rosters %>%
  left_join(cost, by = c("position", "roster.name"))
```

-->



# Check team cost

```{r}


res <- df_rosters %>%
  mutate(tot_cost = cnt * cost) %>%
  group_by(player_id, team_id, position, coach_name, roster.name, tot_cost) %>% # remove two rows per player (skill stacking)
  summarise() %>%
  group_by(team_id, coach_name, roster.name) %>%
  summarise(team_cost = sum(tot_cost)) %>%
  filter(!is.na(team_cost ))

res %>% filter(team_cost > 1150)

```

# remove bribes for TV >1150


all teams with 50 or 100 extra, remove 1 bribe ("get the ref" kick off result ends up on the match page)

```{r}
teams_get_the_ref <- res %>%
  filter(team_cost > 1150) %>%
  ungroup() %>%
  distinct(team_id) %>%
  pull()
```

reduce nr of bribes for get the ref teams with 1

So there are 22 get the ref teams

```{r}
df_rosters %>%
  ungroup() %>%
  filter(position == "1 bribe") %>%
  summarise(sum(cnt))
```
expect 45 - 22 = 23 bribes left

```{r}
# remove the get the ref bribes

df_rosters <- df_rosters %>%
  mutate(cnt = ifelse(team_id %in% teams_get_the_ref & position == "1 bribe", cnt - 1, cnt))

```


```{r}
df_rosters %>%
  ungroup() %>%
  filter(position == "1 bribe") %>%
  summarise(sum(cnt))
```
check!

drop rows with cnt 0

```{r}
df_rosters <- df_rosters %>% 
  filter(cnt != 0)
```


# Check total team cost again

Expect all teams at max 1150 or slightly below.

```{r}


res <- df_rosters %>%
  mutate(tot_cost = cnt * cost) %>%
  group_by(player_id, team_id, position, coach_name, roster.name, tot_cost) %>% # remove two rows per player (skill stacking)
  summarise() %>%
  group_by(team_id, coach_name, roster.name) %>%
  summarise(team_cost = sum(tot_cost)) %>%
  filter(!is.na(team_cost ))

summary(res$team_cost)

```

```{r}
res %>% filter(team_cost < 1135)
```

<!--
```{r}
df_rosters <- df_rosters %>%
  left_join(skill_colors, by = "skill_name")
```

-->

# Add Malta final standing

```{r}
malta_standings <- read_ods(path = "road_to_malta_standings.ods", skip = 2)

malta_standings <- malta_standings %>%
  filter(!is.na(Rank))
```


Make all names lower case before joining.

Add points in coach name.

```{r}
malta_standings$coach_name <- str_to_title(malta_standings$Coach)
df_rosters$coach_name <- str_to_title(df_rosters$coach_name)

malta_standings <- malta_standings %>% 
  mutate(across('coach_name', str_replace, 'Liam_gallagher', 'Liam'))

df_rosters <- df_rosters %>%
  left_join(malta_standings %>% select(coach_name, Points), by = "coach_name")

df_rosters$coach_name <- paste0(df_rosters$coach_name, "_", df_rosters$Points, "p")
```


# Use bbrosterplots to render plots and make pdf

```{r}
create_rosterplot(df_rosters, group_name = "Road to Malta", tournament_ruleset = "EB22")
```

```{r}
#make_skill_table(df_rosters1, group_name = "Road to Malta", tournament_ruleset = "EB22", return_table = FALSE)
```

```{r}

races <- sort(unique(df_rosters$roster.name))

render_rosterbook(params = list(group_name = "Road to Malta", 
                                tournament_ruleset = "EB22", 
                                cover_image = "eurobowl_malta.png",
                                scale_cover_perc = "33%",                                
                                races = races, 
                                perc = FALSE, 
                                team = FALSE), refresh_rmd = TRUE)
```

