---
title: 'Prepare FUMBBL roster data'
author: "Gertjan Verhoeven"
date: '2026-03-31'
params:
  group_name: "Super League S8"
  tournament_ruleset: "EB26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# I. Loading package and data
## Load packages

```{r}
# Load packages
library(tidyverse)
```


## Read the data

We start with reading in the scraped FUMBBL roster data, this is generated by the `fumbbl_rosters` Jupyter Notebook.


```{r}
group_name <- params$group_name

df_rosters1 <- read.csv(file = "input/df_rosters_super_league_s8.csv")

df_rosters1$group_name <- group_name

# Cf. bb_tournament_settings.ods
tournament_ruleset <- params$tournament_ruleset
```

# II. Data cleaning
## Beginning of data cleaning

```{r}
df_rosters <- df_rosters1 %>%
  filter(position != "") %>%
  mutate(cnt = 1) %>%
  rename(skill_name = name)
```



## Adding info on inducements etc
```{r}
# add xtra stuff
extra_stuff <- unique(df_rosters %>% 
                    mutate(apothecary = ifelse(apothecary == "Yes", 1, 0)) %>%
  select(group_name, team_id, roster.name, coach_name, rerolls, assistantCoaches, cheerleaders, apothecary) %>%
    pivot_longer(cols = -c(group_name, team_id, roster.name, coach_name), values_to = "cnt") %>%
    filter(cnt > 0) %>%
    mutate(position = name) %>%
    select(group_name, team_id, roster.name, coach_name, position, cnt) %>%
    mutate(player_id = 99999999) %>%
    mutate(number = 99) %>%
    mutate(skill_id = NA) %>%
    mutate(name = NA)) %>%
    rename(skill_name = name) 

df_rosters <- df_rosters %>%
  select(-c(X, rerolls, assistantCoaches, cheerleaders, apothecary))

df_rosters <- df_rosters %>%
  select(-c(team_name, tournament_id, tournament_name))

df_rosters <- rbind(df_rosters, extra_stuff)

df_rosters <- df_rosters %>%
  mutate(skill_name = if_else(skill_name == "NA", "", skill_name))
```


## Check: skill dist

```{r}
df_rosters %>% 
  filter(!(skill_id %in% c(NA, -1))) %>%
  group_by(skill_id, skill_name) %>%
  summarise(cnt = n()) %>%
  arrange(-cnt)
```


## Check: rosters without skills

```{r}
df_rosters %>%
  mutate(w_skill = ifelse(skill_name %in% c("", NA), 0, 1)) %>%
  group_by(group_name, team_id) %>%
  summarise(n_skills = sum(w_skill)) %>%
  mutate(has_skills = ifelse(n_skills == 0, 0, 1)) %>%
  group_by(group_name) %>%
  summarise(perc_w_skills = mean(has_skills)*100)
```

```{r}
# fix wrong inducement

df_rosters <- df_rosters %>%
  filter(!(position %in% c("1 part-time assistant coach", "1 temp agency cheerleader", "the White Dwarf")))

# drop the white dwarf (we already have Grombrindal :-)


```

## condense skill stacking rows

A player with two skills has two rows, collapse into one, concatenating the two skills.
We sort by skill_id to avoid mirror combinations.

```{r}
df_rosters <- df_rosters %>%
  arrange(player_id, -skill_id) %>%
  group_by(across(c(-skill_id, -skill_name))) %>%
  summarise(skill_name = paste(skill_name, collapse = ""))

df_rosters <- df_rosters %>%
  mutate(skill_name = if_else(skill_name == "NA", "", skill_name))
```

## Check: bribes

Cannot easily seperate out the GEt the ref bribes from the  roster bribes

```{r}
df_rosters %>% 
  ungroup() %>%
  filter(grepl("bribe", position)) %>%
  distinct(position) %>%
  pull()

```

## wrangle bribes into cnt 0-3 bribes

```{r}
df_rosters <- df_rosters %>%
  mutate(cnt = ifelse(grepl("bribe", position), as.integer(substr(position, 1, 2)), cnt)) %>%
  mutate(position = ifelse(grepl("bribes", position), "1 bribe", position)) 
```

# Use bbrosterplots function to add cost

```{r}
library(bbrosterplots)

df_rosters <- df_rosters %>% select(-player_id) %>% ungroup() # PM BUGFIX
df_rosters <- bbrosterplots::add_roster_metadata(df_rosters, edition_val = "bb2025")
```

# Check: roster popularity

```{r}
res <- df_rosters %>%
  group_by(coach_name, roster.name) %>%
  summarise(cnt = n()) %>%
  group_by(roster.name) %>%
  summarise(cnt = n()) %>%
  arrange(-cnt)

gp <- ggplot(res[1:10,], aes(x = reorder(roster.name, cnt) ,y = cnt)) +
  geom_point() +
  expand_limits(y = 0) +
  coord_flip() +
  ylab("# rosters") +
  xlab("") +
  ggtitle("SL8 roster popularity")

ggsave(gp, filename = "roster_pop_sl8.png")

gp
```

### Check: Detect get the ref bribes

Expect all teams at max 1150 or slightly below.

```{r}
res <- df_rosters %>%
  mutate(tot_cost = cnt * cost) %>%
  group_by(player_id, team_id, position, coach_name, roster.name, tot_cost) %>% # remove two rows per player (skill stacking)
  summarise() %>%
  group_by(team_id, coach_name, roster.name) %>%
  summarise(team_cost = sum(tot_cost)) %>%
  filter(!is.na(team_cost )) %>%
  ungroup() %>%
  group_by(roster.name) %>%
  mutate(med_cost = median(team_cost)) %>%
  mutate(n_rosters = n()) %>%
  mutate(delta = team_cost - med_cost)

summary(res$team_cost)
```

Check rosters that do not use their full budget

```{r}
res %>% arrange(roster.name, team_cost, med_cost) %>%
  filter(delta < -40 & n_rosters > 2) 
```

algorithm:
check deviation within roster.
use more than 40k diff.
use only when n > 2

```{r}
get_the_ref_team_ids <- res %>% arrange(roster.name, team_cost, med_cost) %>%
  filter(delta > 40 & n_rosters > 1) %>% pull(team_id)

get_the_ref_team_ids
```

# Remove get the ref bribes

```{r}
df_rosters <- df_rosters %>%
  mutate(cnt = ifelse(team_id %in% get_the_ref_team_ids & position == "1 bribe", cnt - 1, cnt)) %>%
  filter(cnt != 0)
```



# Use bbrosterplots to render plots and make pdf


```{r}
unique(df_rosters$roster.name)

create_rosterplot(df_rosters, group_name = "Super league S8", tournament_ruleset = "EB26")
```

```{r}
#make_skill_table(df_rosters1, group_name = "Super league S7", tournament_ruleset = "BAABowl", return_table = FALSE)
```

```{r}
#res <- file.copy(from = system.file("extdata", "EB25.jpg", package = "bbrosterplots"), to = ".")

races <- sort(unique(df_rosters$roster.name))

render_rosterbook(params = list(group_name = "Super League S8", 
                                tournament_ruleset = "EB26", 
                                cover_image = "super_league_logo.png",
                                scale_cover_perc = "33%",                                
                                races = races, 
                                perc = FALSE, 
                                team = FALSE), refresh_rmd = TRUE)
```
