---
title: "FUMBBL win rates"
author: "Gertjan Verhoeven"
date: '2023-06-03'
output: pdf_document
---

PM pivot naar 1 row per race
Verhaaltje over de snots.

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
```

# Read in data

```{r}
df_matches <- read.csv(file = "../datasets/current/df_matches.csv")
df_matches$match_date <- as.Date(df_matches$match_date)
df_matches$week_date <- as.Date(df_matches$week_date)
```

```{r}
df_mbt <- read.csv(file = "../datasets/current/df_mbt.csv")
df_mbt$match_date <- as.Date(df_mbt$match_date)
df_mbt$week_date <- as.Date(df_mbt$week_date)

df_mbt$raw_tv_bins <- cut(df_mbt$team_value, breaks = c(-Inf, 500, 900, 1100, Inf))
df_mbt$match_tv_bins <- cut(df_mbt$tv_match, breaks = c(-Inf, 1150, 1450, Inf))

df_mbt[df_mbt$race_name == "Khorne",]$bb2020_nov21_tier <- 2
df_mbt[df_mbt$race_name == "Khorne",]$bb2020_tier <- 2
```

```{r}
inducements <- read.csv(file = "../datasets/current/inducements.csv")
```

# Check current tv bin brackets

```{r fig.width = 10}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  select(race_name, tv_match)

ggplot(res, aes(x = tv_match)) + geom_histogram(binwidth = 100) + geom_vline(xintercept = c(1150, 1450), col = "red") 
```
Looks ok.

# Straight up win rates to compare with the FUMBBL stats:

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  group_by(race_name, race_type, bb2020_nov21_tier) %>%
  summarise(n_games = n(),
            win_rate = mean(wins))

ggplot(res, aes(x = reorder(race_name, win_rate), y = win_rate, col = factor(bb2020_nov21_tier))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates ") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```

https://fumbbl.com/p/rosterstats
manual eyeball check: we can reproduce the rosterstats.

```{r}
res_list <- list(main_result = res %>% arrange(-win_rate))
res
```



# Win rates with or without star players

We start with plotting the distribution of team strength (TV) in the BB2020 Competitive division:

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  group_by(race_name, race_type, bb2020_tier, has_sp) %>%
  summarise(n_games = n(),
            win_rate = mean(wins))

ggplot(res, aes(x = reorder(race_name, win_rate), y = win_rate, col = race_type, shape = factor(has_sp))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (with or without star player)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```
```{r}
res_list <- append(res_list, 
                   list(by_starplayer = res %>% pivot_wider(names_from = has_sp, values_from = c(n_games, win_rate))))
```


# main analysis: win rate by team value bracket

First without SP. Select only bins that have at least 300 games in it.

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 0) %>%
  group_by(race_name, race_type, bb2020_nov21_tier, match_tv_bins) %>%
  summarise(n_games = n(),
            win_rate = mean(wins)) 

ggplot(res %>% filter(n_games > 300), aes(x = reorder(race_name, win_rate), y = win_rate, col = factor(bb2020_nov21_tier), shape = factor(match_tv_bins))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (by tv bin, no star players)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```
```{r}
res_list <- append(res_list, 
                   list(by_bracket_no_sp = res %>% arrange(-win_rate)))
```


Next for matches that include star players.

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 1) %>%
  group_by(race_name, race_type, bb2020_nov21_tier, match_tv_bins) %>%
  summarise(n_games = n(),
            win_rate = mean(wins))

ggplot(res %>% filter(n_games > 300), aes(x = reorder(race_name, win_rate), y = win_rate, col = factor(bb2020_nov21_tier), shape = factor(match_tv_bins))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (by tv bin, with at least 1 star player in the match)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```
```{r}
res_list <- append(res_list, 
                   list(by_bracket_with_sp = res %>% arrange(-win_rate)))
```



# What up with the snotlings and star players?

First we check the star player matches:

```{r}
match_ids <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 1 & race_name == "Snotling") %>%
  pull(match_id)
```

Check inducements for these matches:

```{r}
res <- inducements %>% filter(match_id %in% match_ids) %>% arrange(match_id) %>%
  group_by(inducements) %>%
  summarise(cnt = n()) %>% 
  arrange(-cnt)
```

So Hakflem, Bomber and Morg, thats wassup.

```{r}
res_list <- append(res_list, 
                   list(snotling_sp_inducements = res %>% arrange(-cnt)))
```


# What up with the snotlings without star players?

They still have quite high win rate for a stunty team.

Want to know more about snotlings without star player at low TV value before inducement.

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 0) %>%
  group_by(race_name, race_type, bb2020_tier, raw_tv_bins) %>%
  summarise(n_games = n(),
            win_rate = mean(wins)) 

ggplot(res %>% filter(n_games > 100), aes(x = reorder(race_name, win_rate), y = win_rate, col = race_type, shape = factor(raw_tv_bins))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (by RAW tv bin)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```

```{r}
res_list <- append(res_list, 
                   list(main_result_by_bracket_no_sp = res %>% arrange(-win_rate)))
```

So for high team value before inducement they can get pretty strong?

Check what inducements are used in matches where snotlings win.

```{r}
match_ids <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 0 & race_name == "Snotling" & raw_tv_bins == "(1.1e+03, Inf]" & wins == 1) %>%
  pull(match_id)

res <- inducements %>% filter(match_id %in% match_ids) %>% arrange(match_id) %>%
  group_by(inducements) %>%
  summarise(cnt = n()) %>% 
  arrange(-cnt)

res
```
This is riotous rookie.

```{r}
res_list <- append(res_list, 
                   list(snotling_no_sp_inducements = res %>% arrange(-cnt)))
```

```{r eval = FALSE}
df_matches %>% filter(match_id %in% match_ids & team1_coach_id != 255856 & team2_coach_id != 255856)
```

Eyeball a few teams. A snotling team at 1100 before inducement has a lot of skilled up players (trolls, pump wagons etc).

# Win rates by year

Check changes over time in the win rates:

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  group_by(race_name, race_type, bb2020_tier, year) %>%
  summarise(n_games = n(),
            win_rate = mean(wins))

ggplot(res, aes(x = reorder(race_name, win_rate), y = win_rate, col = race_type, shape = factor(year))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (by year)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```

```{r}
res_list <- append(res_list, 
                   list(win_rates_by_year = res %>% arrange(-win_rate)))
```

# write named list of dfs as as multi sheet excel

```{r}
writexl::write_xlsx(res_list, "FUMBBL_competitive_win_rates_may2023.xlsx")
```