---
title: "FUMBBL win rates"
author: "Gertjan Verhoeven"
date: '2023-06-03'
output: pdf_document
---

PM pivot naar 1 row per race
Verhaaltje over de snots.

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
```

# Read in data

```{r}
df_matches <- read.csv(file = "../datasets/current/df_matches.csv")
df_matches$match_date <- as.Date(df_matches$match_date)
df_matches$week_date <- as.Date(df_matches$week_date)
```

```{r}
df_mbt <- read.csv(file = "../datasets/current/df_mbt.csv")
df_mbt$match_date <- as.Date(df_mbt$match_date)
df_mbt$week_date <- as.Date(df_mbt$week_date)

df_mbt$raw_tv_bins <- cut(df_mbt$team_value, breaks = c(-Inf, 500, 900, 1100, Inf))
df_mbt$team_tv_bins <- cut(df_mbt$team_value, breaks = c(-Inf, 1150, 1450, Inf))
df_mbt$match_tv_bins <- cut(df_mbt$tv_match, breaks = c(-Inf, 1150, 1450, Inf))

df_mbt[df_mbt$race_name == "Khorne",]$bb2020_nov21_tier <- 2
df_mbt[df_mbt$race_name == "Khorne",]$bb2020_tier <- 2
```

```{r}
inducements <- read.csv(file = "../datasets/current/inducements.csv")
```

# Check current tv bin brackets

```{r fig.width = 10}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  select(race_name, team_value)

ggplot(res, aes(x = team_value)) + geom_histogram(binwidth = 100) + geom_vline(xintercept = c(1150, 1450), col = "red") 
```
Looks ok.

# Straight up win rates to compare with the FUMBBL stats:

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  group_by(race_name, race_type, bb2020_nov21_tier) %>%
  summarise(n_games = n(),
            win_rate = mean(wins))

ggplot(res, aes(x = reorder(race_name, win_rate), y = win_rate, col = factor(bb2020_nov21_tier))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates ") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```

https://fumbbl.com/p/rosterstats
manual eyeball check: we can reproduce the rosterstats.

```{r}
res_list <- list(main_result = res %>% arrange(-win_rate))
res
```



# Win rates with or without star players

We start with plotting the distribution of team strength (TV) in the BB2020 Competitive division:

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  group_by(race_name, race_type, bb2020_tier, has_sp) %>%
  summarise(n_games = n(),
            win_rate = mean(wins))

ggplot(res, aes(x = reorder(race_name, win_rate), y = win_rate, col = race_type, shape = factor(has_sp))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (with or without star player)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```
```{r}
res_list <- append(res_list, 
                   list(by_starplayer = res %>% pivot_wider(names_from = has_sp, values_from = c(n_games, win_rate))))
```


# main analysis: win rate by team value bracket

We focus on team value, not match value, i.e. before inducement.
This focusses on how developed the players / team is.

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  group_by(race_name, race_type, bb2020_nov21_tier, team_tv_bins) %>%
  summarise(n_games = n(),
            win_rate = mean(wins)) 

ggplot(res %>% filter(n_games > 300), aes(x = reorder(race_name, win_rate), y = win_rate, col = factor(bb2020_nov21_tier), shape = factor(team_tv_bins))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (by tv bin, including SP)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```
Select only bins that have at least 300 games in it.

```{r}
res_list <- append(res_list, 
                   list(main_result_by_tv_bracket = res %>% 
                          filter(n_games > 300) %>% 
                          pivot_wider(names_from = team_tv_bins, values_from = c(n_games, win_rate)) %>% 
                          arrange(race_name))
                   )
```

Now without SP. 

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 0) %>%
  group_by(race_name, race_type, bb2020_nov21_tier, match_tv_bins) %>%
  summarise(n_games = n(),
            win_rate = mean(wins)) 

ggplot(res %>% filter(n_games > 300), aes(x = reorder(race_name, win_rate), y = win_rate, col = factor(bb2020_nov21_tier), shape = factor(match_tv_bins))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (by tv bin, no star players)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```
```{r}
res_list <- append(res_list, 
                   list(by_bracket_no_sp = res %>% arrange(-win_rate)))
```


Next for matches that include star players.

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 1) %>%
  group_by(race_name, race_type, bb2020_nov21_tier, match_tv_bins) %>%
  summarise(n_games = n(),
            win_rate = mean(wins))

ggplot(res %>% filter(n_games > 300), aes(x = reorder(race_name, win_rate), y = win_rate, col = factor(bb2020_nov21_tier), shape = factor(match_tv_bins))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (by tv bin, with at least 1 star player in the match)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```
```{r}
res_list <- append(res_list, 
                   list(by_bracket_with_sp = res %>% arrange(-win_rate)))
```



# What up with the snotlings and star players?

First we check the star player matches:

```{r}
match_ids <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 1 & race_name == "Snotling") %>%
  pull(match_id)
```

Check inducements for these matches:

```{r}
res <- inducements %>% filter(match_id %in% match_ids) %>% arrange(match_id) %>%
  group_by(inducements) %>%
  summarise(cnt = n()) %>% 
  arrange(-cnt)

res
```

So Hakflem, Bomber and Morg, thats wassup.

```{r}
res_list <- append(res_list, 
                   list(snotling_sp_inducements = res %>% arrange(-cnt)))
```


# What up with the snotlings without star players?

They still have quite high win rate for a stunty team.

Want to know more about snotlings without star player at low TV value before inducement.

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 0) %>%
  group_by(race_name, race_type, bb2020_tier, raw_tv_bins) %>%
  summarise(n_games = n(),
            win_rate = mean(wins)) 

ggplot(res %>% filter(n_games > 100), aes(x = reorder(race_name, win_rate), y = win_rate, col = race_type, shape = factor(raw_tv_bins))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (no sp, by RAW tv bin)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```

```{r}
res_list <- append(res_list, 
                   list(main_result_by_bracket_no_sp = res %>% arrange(-win_rate)))
```

So for high team value before inducement they can get pretty strong? This means that they have a lot of skills.

# Check snotling inducements

Check what inducements are used in matches where snotlings win.

```{r}
match_ids <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 0 & race_name == "Snotling" & wins == 1) %>%
  pull(match_id)

res <- inducements %>% filter(match_id %in% match_ids) %>% arrange(match_id) %>%
  group_by(inducements) %>%
  summarise(cnt = n()) %>% 
  arrange(-cnt)

res
```
This is riotous rookie.

```{r}
res_list <- append(res_list, 
                   list(snotling_no_sp_inducements = res %>% arrange(-cnt)))
```

```{r eval = FALSE}
df_matches %>% filter(match_id %in% match_ids & team1_coach_id != 255856 & team2_coach_id != 255856)
```

Eyeball a few teams. A snotling team at 1100 before inducement has a lot of skilled up players (trolls, pump wagons etc).

```{r}
tmp <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 0 & race_name == "Snotling" )
```

```{r}
tmp <- df_mbt %>%
  filter(division_name == "Competitive" & has_sp == 0 & race_name == "Snotling" & team_value > 750 & team_value < 900)
```


# Win rates by year

Check changes over time in the win rates:

```{r fig.height= 8, fig.width = 8}
res <- df_mbt %>%
  filter(division_name == "Competitive") %>%
  group_by(race_name, race_type, bb2020_tier, year) %>%
  summarise(n_games = n(),
            win_rate = mean(wins))

ggplot(res, aes(x = reorder(race_name, win_rate), y = win_rate, col = race_type, shape = factor(year))) +
  geom_point(aes(size = n_games)) + coord_flip() + expand_limits(y = 0) +
  scale_size_area() + ggtitle("FUMBBL Competitive division win rates \n (by year)") +
  labs(y = "Win rate (%)", x = "") + geom_hline( yintercept =  0.5, col = "black")
```

```{r}
res_list <- append(res_list, 
                   list(win_rates_by_year = res %>% arrange(-win_rate)))
```

# write named list of dfs as as multi sheet excel

```{r}
selection_of_tables <- c("main_result", "main_result_by_tv_bracket", "by_starplayer")

writexl::write_xlsx(res_list[selection_of_tables], "FUMBBL_competitive_win_rates_may2023.xlsx")
```

# Snotling analysis

Snotlings are a stunty team, but have a surprisingly high win rate of 54%.
Snotling teams achieve this result by keeping team value low, and then inducing famous Star Players (so called "Mega stars"): mostly Morg n Thorg and Bomber Dribblesnot, and to a lesser extent Hakflem, Fungus and Varag.

However, if we select only Snotling matches that did not include a Star player, the win rate drops to 46%, which is  higher than other "non-stunty" teams such as Black Orcs or Old World Alliance. Most of those matches were with team values (before inducement) around 850-1050K, which for a Snotling team typically means a lot of skilled up players: Trolls, Pump wagons, multiple sneaky git (SG), etc. Riotous Rookie and bribes complete the package. Finally, these teams are more likely to be played by experienced coaches, that this analysis does not take into account. The recent nerf to SG might lower this rate again in the future. 


